---
title: "L03 ggplot II"
author: "Data Visualization (STAT 302)"
date: "David Cheng"
output:
  html_document:
    toc: true
    toc_float: true
---

# Overview

The goal of this lab is to continue the process of unlocking the power of `ggplot2` through constructing and experimenting with a few basic plots.

# Datasets

We'll be using data from the `BA_degrees.rda` and `dow_jones_industrial.rda` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. Below is a description of the variables contained in each dataset.

`BA_degrees.rda`

* `field` - field of study
* `year_str` - academic year (e.g. 1970-71)
* `year` - closing year of academic year
* `count` - number of degrees conferred within a field for the year
* `perc` - field's percentage of degrees conferred for the year

`dow_jones_industrial.rda`

* `date` - date
* `open` - Dow Jones Industrial Average at open
* `high` - Day's high for the Dow Jones Industrial Average
* `low` - Day's low for the Dow Jones Industrial Average
* `close` - Dow Jones Industrial Average at close
* `volume` - number of trades for the day

We'll also be using a subset of the BRFSS (Behavioral Risk Factor Surveillance System) survey collected annually by the Centers for Disease Control and Prevention (CDC). The data can be found in the provided `cdc.txt` file --- place this file in your `/data` subdirectory. The dataset contains 20,000 complete observations/records of 9 variables/fields, described below.

* `genhlth` - How would you rate your general health? (excellent, very good, good, fair, poor)
* `exerany` - Have you exercised in the past month? (`1 = yes`, `0 = no`) 
* `hlthplan` - Do you have some form of health coverage? (`1 = yes`, `0 = no`) 
* `smoke100` - Have you smoked at least 100 cigarettes in your life time? (`1 = yes`, `0 = no`) 
* `height` -  height in inches
* `weight` - weight in pounds
* `wtdesire` - weight desired in pounds
* `age` - in years
* `gender` - `m` for males and `f` for females

```{r load-packages, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(lubridate)
library(splines)
```

# Exercises

Complete the following exercises.

<br>

### Exercise 1

Using `BA_degrees` dataset, recreate the following graphics as precisely as possible.

```{r}
load(file = "data/BA_degrees.rda")

# Wrangling for plotting
ba_dat <- BA_degrees %>%
  # mean % per field
  group_by(field) %>% 
  mutate(mean_perc = mean(perc)) %>% 
  # Only fields with mean >= 5%
  filter(mean_perc >= 0.05) %>%
  # Organizing for plotting
  arrange(desc(mean_perc), year) %>% 
  ungroup() %>% 
  mutate(field = fct_inorder(field))
```

<br>

##### Plot 1
```{r}
ba_dat %>% 
  ggplot(aes(x = year, y = perc)) +
  geom_line() +
  facet_wrap(~field) +
  labs(x = "Year",
       y = "Proportion of Degrees")
```
<br>

##### Plot 2
```{r}
ba_dat %>% 
  ggplot(aes(x = year, y = perc, color = "red")) +
  geom_line() +
  geom_area(fill = "red", alpha = .5) +
  facet_wrap(~field) +
  labs(x = "Year",
       y = "Proportion of Degrees")
```
<br>

##### Plot 3
```{r}
ba_dat %>% 
  ggplot(aes(x = year, y = perc, color = field)) +
  geom_line() +
  labs(x = "Year",
       y = "Proportion of Degrees")
```


<br>

### Exercise 2

Using `dow_jones_industrial` dataset, recreate the following graphics as precisely as possible.

```{r}
load(file = "data/dow_jones_industrial.rda")

# Restrict data to useful range
djia_date_range <- dow_jones_industrial %>% 
  filter(date >= ymd("2008/12/31") & date <= ymd("2010/01/10")) 
```

<br>

##### Plot 1
```{r}
djia_date_range %>% 
  ggplot(aes(x = date, y = close)) +
   geom_line(color = 'purple') +
   geom_smooth(color = 'green', fill = 'red', width = .75)
```
<br>

##### Plot 2
```{r}
djia_date_range %>% 
  ggplot(aes(x = date, y = close)) +
    geom_line() +
    geom_smooth(span = 0.3, se = FALSE)
```
<br>

##### Plot 3
```{r}
djia_date_range %>% 
  ggplot(aes(x = date, y = close)) +
    geom_line() +
    geom_smooth(method = lm, formula = y ~ ns(x,6), se = FALSE)
```
<br>

### Exercise 3

Using `cdc` dataset, recreate the following graphics as precisely as possible.

```{r, warning=FALSE, message=FALSE}
# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))
```

<br>

##### Plot 1

Consturct this plot in two ways. Once using `cdc` and once using the `genhlth_count`

```{r}
genhlth_count <- cdc %>% 
  count(genhlth)

cdc %>% 
  ggplot(aes(x = genhlth)) +
  geom_bar()

genhlth_count %>% 
  ggplot(aes(x = genhlth, y = n)) +
  geom_col()
```

<br>

##### Plot 2
```{r}
cdc2 <- cdc %>% 
  mutate(hlthplan = as.factor(hlthplan))

cdc2 %>% 
  ggplot(aes(x = genhlth, fill = hlthplan)) +
  geom_bar(position = "dodge")
```
<br>

##### Plot 3
```{r}
cdc %>% 
  ggplot(aes(x = weight, color = genhlth, fill = genhlth)) +
  geom_density(alpha = .2) +
  facet_wrap(~ gender)
```
<br>

##### Plot 4
```{r}
cdc %>% 
  ggplot(aes(x = weight, color = gender, fill = gender)) +
  geom_density(alpha = .5) +
  facet_wrap(~ genhlth, ncol = 2) +
  xlim(c(50, 300))
```
<br>

##### Plot 5
```{r}
cdc %>% 
  ggplot(aes(x = gender, y = height, fill = gender)) +
  geom_boxplot(alpha = .4) +
  facet_wrap(~genhlth, nrow = 1)
```
<br>

##### Plot 6
```{r}
cdc %>% 
  ggplot(aes(x = height, y = weight, color = gender)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm", se = FALSE)
```

