---
title: "L01 Introduction"
author: "David Cheng"
date: "April 2, 2019"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load-packages, warning=FALSE, message=FALSE}
# Load package(s) - We will do this together 
library(tidyverse)
library(ggstance)
library(skimr)
library(janitor)
```

# Dataset

We'll be using data from the [**lego** package](https://github.com/seankross/lego) which is already in the `/data` subdirectory, along with many other processed datasets, as part of the zipped folder for this lab. 

### Task 1
Download and install [R Software](https://cran.rstudio.com/).
<br><br>
**Completed**

<br>

### Task 2
1. Download and install [RStudio](https://www.rstudio.com/products/rstudio/download4/).
2. Open RStudio and ensure it and R have been properly installed. 
3. Go to **Tools** > **Global Options** and change the two indicated settings below and click **Apply**. --- See L01_introduction.html.
<br><br>
**Completed**

<br>

### Task 3
Install the following packages:

- `tidyverse`
- `ggstance`
- `skimr`
<br><br>
**Completed**

<br>

### Task 4

1. Download `data_vis_labs.zip` from Canvas. 
2. Unzip the file and place the unzipped `data_vis_labs` directory where you would like to keep all of your lab work for this course.
3. Open RStudio and create a project folder for this *exsisting directory*.  
4. Appropriately rename `template_L01.Rmd` for submission (e.g. *Kuper_Arend_L01.Rmd*).
5. Compile the *_L01.Rmd file with Cmd/Ctrl + Shift + K.
<br><br>
**Completed**


<br>

### Task 5 
Optional: It is always handy to have a versatile text editor and I would suggest downloading [Sublime Text](https://www.sublimetext.com/). It is free to use.
<br><br>
**Completed**
<br>

# Exercise 1

### Inspect the data
The **lego** package provides a helpful dataset some interesting variables. Let's take a quick look at the data.
```{r}
load(file = "data/legosets.rda")

legosets <- legosets %>% 
  clean_names()

skim_with(numeric = list(hist = NULL),
          integer = list(hist = NULL))

legosets %>% 
  skim()
```

### Price per year
First, let's look at the average cost of LEGO sets over time. The main variable of interest here is `USD_MSRP`, or the manufacturer's suggested retail price in constant dollars (i.e. not adjusted for inflation).
```{r}
avg_price_per_year <- legosets %>% 
  filter(!is.na(usd_msrp)) %>% 
  group_by(year) %>%
  summarise(price = mean(usd_msrp))

avg_price_per_year %>% 
  ggplot(aes(x = year, y = price)) +
    geom_point(size = .5) +
    geom_line(color = "#444444", size = .5) +
    labs(x = "Year",
         y = "Price (USD)",
         title = "Average Price of Lego sets",
         subtitle = "Amounts are reported in current USD",
         caption = "Source: Lego") +
    theme_minimal()
```
<br>

> The average price of lego sets goes up over time.

### Pieces per year
Next, let’s look at how the number of pieces per set has changed over time. Because Duplo sets are much smaller (since they’re designed for toddlers), we’ll make a special indicator variable for them.
```{r}
pieces_per_year <- legosets %>% 
  filter(year >= 1986) %>% 
  filter(!is.na(pieces)) %>%
  mutate(duplo = ifelse(theme == "Duplo", "Duplo", "Lego")) %>% 
  group_by(year, duplo) %>%
  summarise(pieces = mean(pieces),
            num_sets = n()) %>% 
  arrange(year)

pieces_per_year %>% 
  ggplot(aes(x = year, y = pieces)) +
    geom_point(size = .5) +
    geom_line(color = "#444444", size = .5) +
    facet_wrap(~duplo) +
    labs(x = "Year",
         y = "Pieces",
         title = "Average Pieces of Lego sets",
         caption = "Source: Lego") +
    theme_minimal()
```
<br>

> Duplo sets hover around 5 to 50 pieces per set with a slow increase in the number of pieces over time. All non Duplo sets have seen an increase in the average number of pieces per set over time

### LEGO set themes
In the 1990s, [LEGO began partnering with famous brands and franchises](https://www.cnbc.com/2018/04/27/lego-marketing-strategy-made-it-world-favorite-toy-brand.html) to boost its own sales. First, let's see how many different "themes" LEGO now offers:
```{r}
legosets %>% 
  distinct(theme)

theme_counts <- legosets %>% 
  count(theme, sort = TRUE) %>% 
  mutate(theme = fct_inorder(theme, ordered = TRUE))

theme_counts %>% 
  filter(n > 150) %>% 
  ggplot(aes(x = theme, y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL,
       y = "Number of Sets")  

theme_counts %>% 
  filter(n > 150) %>% 
  ggplot(aes(x = n, y = fct_rev(theme))) +
  geom_barh(stat = "identity") +
  labs(x = "Number of Sets",
       y = NULL) 
```
<br>

> Duplo has the most number of sets followed by Star Wars, Town, and Coleectable Minifigures.

