---
title: "L09 Themes"
author: "Data Visualization (STAT 302)"
date: "David Cheng"
output:
  html_document:
    toc: true
    toc_float: true
---
 
# Datasets

```{r load-packages-data-seed, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(ggthemes)
# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor"),
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")
  ))

# Set seed
set.seed(8221984)

# Selecting a random subset of size 100
cdc_small <- cdc %>% sample_n(100)
```

<br>

### Exercise 1

Use the `cdc_small` dataset to explore several pre-set `ggthemes`. The code below constructs the familiar scatterplot of `weight` by `height` and stores it in `plot_01`. Display `plot_01` to observe the default theme. **Explore/apply, and display at least 7 other pre-set themes** from the `ggplot2` or `ggthemes` package. Don't worry about making adjustments to the figures under the new themes. Just get a sense of what the themes are doing to the original figure `plot_01`. 

Which theme or themes do you particularly like? Why? There should be at least 8 plots for this task. `temp1` is pictured below.

<br>

```{r,echo=FALSE, out.width = '50%', fig.show='hold'}
# Building plot
plot_01 <- ggplot(data = cdc_small, aes(x = height, y = weight)) +
  geom_point(size = 3, aes(shape = genhlth, color = genhlth)) +
  scale_y_continuous(
    name = "Weight in Pounds",
    limits = c(100, 300),
    breaks = c(seq(100, 350, 25)),
    trans = "log10",
    labels = scales::unit_format(unit = "lbs")
  ) +
  scale_x_continuous(
    name = "Height in Inches",
    limits = c(55, 80),
    breaks = seq(55, 80, 5),
    labels = scales::unit_format(unit = "in")
  ) +
  scale_shape_manual(
    name = "General\nHealth?",
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"),
    values = c(17, 19, 15, 9, 4)
  ) +
  scale_color_brewer(
    name = "General\nHealth?",
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"),
    palette = "Set1"
  ) +
  theme(
    legend.position = c(1, 0),
    legend.justification = c(1, 0)
  ) +
  labs(title = "CDC BRFSS: Weight by Height")

plot_01 +
  theme_base()

plot_01 +
  theme_calc()
```


```{r,echo=FALSE, out.width = '50%', fig.show='hold'}
plot_01 +
  theme_economist()

plot_01 +
  theme_economist_white()
```

```{r,echo=FALSE, out.width = '50%', fig.show='hold'}
plot_01 +
  theme_excel_new()

plot_01 + 
  theme_fivethirtyeight()
```

```{r,echo=FALSE, out.width = '50%', fig.show='hold'}
plot_01 +
  theme_wsj()

plot_01 + 
  theme_stata()
```
<br><br><br>

> The WSJ journal theme looks particularily bad with the weird dotted lines. The color and the font look outdated. I guess they're trying to appear to WSJ readers who are generally older. The Excel and 538 ones are my favorites because they have grid lines and the plots looks clean and sharp.

<br>

### Exercise 2
```{r}
plot_01 +
  theme(plot.title = element_text(color = "#00cc33", hjust = 1, face = "bold", margin = margin()),
        plot.background = element_rect(fill = "#ffff00", colour = "#39ff14", size = 10),
        axis.text = element_text(colour = "#f442aa"),
        axis.title = element_text(colour = "#7741f4"),
        axis.text.x = element_text(angle = -180, vjust = 1, hjust = 0),
        axis.text.y = element_text(angle = -90, vjust = 1, hjust = 0),
        axis.ticks.length = unit(2, "cm"),
        legend.background = element_rect(fill = "#00FFFF", colour = "#FF0000", size = 1),
        panel.grid.major = element_line(color = "#FF0000", size = 0.8),
        panel.grid.minor = element_line(color = "#3EFF00", size = 0.8),
        panel.background = element_rect(fill = "#000100"))
```

<br>


### Exercise 3

```{r}
cdc_small %>% 
  ggplot(aes(height, weight, color = genhlth, shape = genhlth)) +
  geom_point(size = 3) +
  labs(title = "CDC BRFSS: Weight by Height", x = "Height (Inches)", y = "Weight (Pounds)", color = "General\nHealth?", shape = "General\nHealth?") +
  scale_color_manual(values = c("Excellent" = "#008656", "Very Good" = "#7FCECD", 
                                "Good" = "#0D2D6C", "Fair" = "#FFC520", "Poor" = "#EF553F")) +
  scale_shape_manual(values = c("Excellent" = 17, "Very Good" = 19, "Good" = 15, "Fair" = 9, "Poor" = 4)) +
  theme(plot.title = element_text(color = "#4E2A84", face = "bold", margin = margin(t = 10, l = 5, b = 5)),
        plot.background = element_rect(fill = "white", colour = "#4E2A84", size = 3),
        legend.title = element_text(color = "#4E2A84", face = "bold"),
        legend.text = element_text(color = "#4E2A84", face = "bold"),
        axis.text = element_text(colour = "#4E2A84"),
        axis.title = element_text(colour = "#4E2A84", face = "bold"))

cdc_small %>% 
  ggplot(aes(height)) +
  geom_bar(fill = "#836EAA") +
  labs(title = "CDC BRFSS: Distribution of Heights", x = "Height (Inches)", y = "Number of Observations", color = "General\nHealth?", shape = "General\nHealth?") +
  theme(plot.title = element_text(color = "#4E2A84", face = "bold", margin = margin(t = 10, l = 5, b = 5)),
        plot.background = element_rect(fill = "white", colour = "#4E2A84", size = 3),
        axis.text = element_text(colour = "#4E2A84"),
        axis.title = element_text(colour = "#4E2A84", face = "bold"))
```




