---
title: "L5 Toolbox"
author: "Data Visualization (STAT 302)"
date: "David Cheng"
output:
  html_document:
    toc: true
    toc_float: true
---

# Overview

The goal of this lab is to explore more useful plots in `ggplot2`. Specfically we will be focusing on surface plots and geospatial plots (maps).

# Datasets

We'll be using data from the `blue_jays.rda`, `cdc.txt`, and the `US_income.rda` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. 

```{r, message = FALSE}
# Load package(s)
library(tidyverse)
library(purrr)
library(viridis)
library(statebins)
library(sf)
# Load datasets
load("data/blue_jays.rda")
load("data/US_income.rda")

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))
```

### Exercise 1

Using `blue_jays.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- `geom_density_2d()` or `stat_density_2d()`
- Transparency is 1/3
- Horizontal lower limit is 57 and upper limit is 82
- Point size 1.5
- Line size is 0.4
- `binwidth` set to 0.004
- Minimal theme
```{r}
blue_jays %>% 
  ggplot(aes(Mass, Head)) +
  geom_point(size = 1.5, alpha = 1/3) +
  xlim(c(57, 82)) +
  # stat_density_2d() +
  geom_density_2d(size = 0.4, color = "black", binwidth = 0.004) +
  theme_minimal() +
  labs(x = "Body mass (g)", y = "Head length (mm)")
```
<br>

### Exercise 2

Using `cdc` dataset, recreate the following graphics as precisely as possible.

##### Plot 1
```{r}
cdc %>% 
  ggplot(aes(height, weight)) +
  geom_hex(bins = 35) + 
  labs(x = "Height (in)", y = "Weight (lbs)")
```

<br>

##### Plot 2
```{r}
cdc %>% 
  ggplot(aes(height, weight)) +
  stat_density_2d(aes(colour = ..level.., fill = ..level..), geom = "polygon", show.legend = FALSE) +
  facet_wrap(~ gender) +
  theme_minimal() + 
  labs(x = "Height (in)", y = "Weight (lbs)")
```
<br>

### Exercise 3
```{r, message = FALSE}
il_counties <- map_data("county", "illinois") %>%
  select(lon = long, lat, group, id = subregion)

ggplot(il_counties, aes(lon, lat)) +
  geom_polygon(aes(group = group), fill = NA, colour = "grey50") + 
  coord_quickmap() +
  theme_void() +
  labs(title = "Illinois")
```
<br>

### Exercise 4

Using `US_income` dataset, recreate the following graphics as precisely as possible.

<br>

```{r}
# Setting income levels
US_income <- mutate(
  US_income,
  income_bins = cut(
    ifelse(is.na(median_income), 25000, median_income),
    breaks = c(0, 40000, 50000, 60000, 70000, 80000),
    labels = c("< $40k", "$40k to $50k", "$50k to $60k", "$60k to $70k", "> $70k"),
    right = FALSE
  )
)
```

<br>


##### Plot 1
```{r}
US_income %>% 
  ggplot(aes(geometry = geometry, fill = income_bins)) +
  geom_sf(size = 0.2, color = "grey80") +
  scale_fill_viridis(discrete = TRUE) +
  labs(fill = "Median\nIncome") +
  theme_void() +
  coord_sf(datum = NA)
```
<br>

##### Plot 2
```{r}
US_income %>% 
  ggplot(aes(state = name, fill = income_bins)) +
  geom_statebins() +
  scale_fill_viridis(discrete = TRUE) +
  labs(fill = "Median\nIncome") +
  theme_statebins() 
```
<br>


